---
title: "HW: Creating Vectors"
author: "PUT YOUR NAME HERE"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: cosmo
    highlight: textmate
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

Collaborators: 

***

### Problem 1

Create vectors to represent the set of possible outcomes for the following scenarios using `:`, `c()`, `seq()`, and/or `rep()`. Save each vector using the provided object name and surround the entire statement with `()` to print out the results. 

```{r}
(example <- c("store", "and", "print", "a", "vector", "like", "this"))
```

a) The sample space of a random variable X is {0, 1, 2, ..., 10}.

```{r}
v1 <- 
```

b) Scores on a test can be 0 or any positive, even integer up to 100.

```{r}
v2 <- 
```

c) A machine reads the pH of a solution to one decimal place of precision.
 
```{r}
v3 <- 
```

d) A player in a particular game can choose rock, paper, scissors, lizard, or Spock.

```{r}
v4 <- 
```

e) A clock counts down from 60 seconds to zero seconds in whole-second increments.

```{r}
v5 <- 
```

f) A particular song's chorus contains the following three phrases, sung once in order and then repeated.
    + I get knocked down
    + But I get up again
    + You're never gonna keep me down
    
```{r}
v6 <- 
```

g) In a group of 100 people, each person has one of four possible blood types: A, B, AB, or O. (Use the `rep` function.)
    + A = 40% of the people
    + B = 11% of the people
    + AB = 4% of the people
    + O = 45% of the people
    
```{r}
v7 <- 
```

### Problem 2

Create a list from all the vectors you created in Problem 1. Give each element in the list (i.e., vector) a name of your choice.

```{r}
list1 <- 
```

```{r}
str(list1)
```

### Problem 3

Consider the following example of a [factorial experiment](https://en.wikipedia.org/wiki/Factorial_experiment) on Wikipedia: An engineer wants to increase the filtration rate of a process and reduce the amount of formaldehyde used. Four factors were considered: temperature (A), pressure (B), formaldehyde concentration (C), and stirring rate (D). Each of the four factors was tested at two levels. The minus (−) and plus (+) signs in the table indicate whether the factor was run at a low (-) or high (+) level. Create a tibble of the "Design matrix and resulting filtration rate" table showing variables A, B, C, D and the resulting filtration rate. Use "low" and "high" as category values rather than `+` and `-` symbols, since those indicate mathematical operations in `R`. Create each variable vector as efficiently as possible using the vector creation functions you have learned.

```{r}
library(tidyverse)
experiment <-
```

```{r}
glimpse(experiment)
```

***
```{r}
Sys.info()
```
