---
title: "Worksheet: Introduction to Datasets"
subtitle: "STT 2860 Intro to Data Management & Visualization" 
author: "PUT YOUR NAME HERE"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: true
    toc_depth: 4 
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

***

* [Dr. Thomley's Introduction to Datasets slides](https://stat-jet-asu.github.io/STT2860DataScience1/Slides/IntroDatasets.html#1)

***

### Packages Used

```{r packagesused}
# load all packages used here
library(HistData)
library(tidyverse) # this loads 9 different packages
library(nycflights13)
```


### The `Nightingale` Dataset

The `Nightingale` dataset is found in **HistData** package.

*erase this and add your own notes*

#### create an object

```{r}
nightingale <- Nightingale
```

#### explore with `head()`

```{r}
head(nightingale)        # default head, 6 rows
head(nightingale, n = 2) # specify number of rows
```

#### explore with `tail()`

```{r}
tail(nightingale)
```

#### explore with `str()`

```{r}
str(nightingale)  # this is what we see in Environment tab
```

#### explore with `glimpse()`

This is the tidyverse version of `str()`. It contains mostly the same information, but more neatly arranges to scale to the window.

```{r}
glimpse(nightingale)
```

#### explore with other functions

This function tells us the row names (or numbers) and column names, as well as the class of the object. Class determines how R interacts with the object.

```{r}
attributes(nightingale)
```

This tells us the names of the columns / variables.

```{r}
names(nightingale)
```

These functions tell us the dimensions of the rectangular dataset, row first and then column for `dim()`.

```{r}
dim(nightingale)  # always row, column
nrow(nightingale)
ncol(nightingale)
```

This tells us the class of the object, like in `attributes()`.

```{r}
class(nightingale)
```

This tells us how R stores the object.

```{r}
typeof(nightingale)
```


### The `flights` Dataset

The `flights` dataset is found in **nycflights13** package.

The nycflight13 package contains 5 related datasets about flights that flew out of NYC airports in 2013.

#### What `class()` is the flights dataset?

The dataset flights has 3 classes. It is a data.frame, but it is also a tibble. A tibble is an 'upgraded' version of a dataframe. For a lot of what we do, it doesn't make a difference,but it can in more advanced programming.

```{r}
class(flights)
```

#### What `typeof()` is the flights dataset?

The data.frame and tibble classes are special subsets of a structure called a list, so that's how they're stored. This is more important to programmers than to what we'll be doing in this class.

```{r}
typeof(flights)
```

#### How many observations (rows) are there?

```{r}
dim(flights)
nrow(flights)
```

#### What are the names of the variables?

```{r}
names(flights)
head(flights)
```

#### How many complete cases are there? 

Cases is synonymous with observations (or rows) here.

```{r}
nrow(flights)
```

#### How many incomplete cases are there? 

Input is a data.frame or tibble, output is a logic vector with a true/false for every row. 

```{r}
flights_complete <- complete.cases(flights)
```

We can add a logic vector since TRUE = 1 and FALSE = 0, so this will tell me how many complete cases I have.

```{r}
sum(complete.cases(flights))
```

The exclamation point (!) in R means "not". 

```{r}
!complete.cases(flights)
```

This tells me how many incomplete cases I have. This means at least one variable is missing data. I need to do further explorations to figure out where the missing values are and if there are some patterns.

```{r}
sum(!complete.cases(flights))
```

### Building from Scratch

*erase this and add your own notes*

#### build a data.frame

Each row in this code will become one column or variable in my dataset. I am creating 4 vectors and connecting them with the `data.frame()` function. All the vectors have to be the same length.

```{r}
dataset1 <- data.frame(
  ID = 1:5,               # variable name = variable contents
  gender = c("male", "male", "female", "female", "nonbinary"),
  height = c(71, 73, 64, 64, 66),
  weight = c(175, 225, 130, 125, 165)
)

dataset1
class(dataset1)
```

#### build a tibble

```{r}
dataset2 <- tibble(
  ID = 1:5,
  gender = c("male", "male", "female", "female", "nonbinary"),
  height = c(71, 73, 64, 64, 66),
  weight = c(175, 225, 130, 125, 165)
)

dataset2
```

#### build a list

```{r}
listex <- list(
  ID = 1:5,
  gender = c("male", "male", "female", "female", "nonbinary"),
  height = c(71, 73, 64, 64, 66),
  weight = c(175, 225, 130, 125, 165)
)

listex
```

```{r}
list1 <- list(
  a = 2,                # constant/vector of length 1, named a
  b = 1:3,              # integer vector of length 3, named b
  d = c("Fred", "Joe")  # character vector of length 2, named d
)

list1

list1$a
```

```{r}
list2 <- list(
  dataA = dataset1,   # dataset we made earlier, named dataA
  dataB = dataset2,   # dataset we made earlier, named dataB
  dataC = mtcars      # dataset included in R, named dataC
)

list2
```


***
```{r}
sessionInfo()
```

