---
title: "Worksheet: Introduction to Datasets"
subtitle: "STT 2860 Intro to Data Management & Visualization" 
author: "PUT YOUR NAME HERE"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: true
    toc_depth: 4 
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

***

* [Dr. Thomley's Introduction to Datasets slides](https://stat-jet-asu.github.io/STT2860DataScience1/Slides/IntroDatasets.html#1)

***

### Packages Used

```{r packagesused}
# load all packages used here
library(HistData)
```


### The `Nightingale` Dataset

The `Nightingale` dataset is found in **HistData** package.

*erase this and add your own notes*

#### create an object

```{r}
nightingale <- Nightingale
```

#### explore with `head()`

```{r}
head(nightingale)        # default head, 6 rows
head(nightingale, n = 2) # specify number of rows
```

#### explore with `tail()`

```{r}
tail(nightingale)
```

#### explore with `str()`

```{r}
str(nightingale)  # this is what we see in Environment tab
```

#### explore with `glimpse()`

```{r}

```

#### explore with other functions

```{r}

```


### The `flights` Dataset

The `flights` dataset is found in **nycflights13** package.

*erase this and add your own notes*

#### What `class()` is the flights dataset?

```{r}

```

#### What `typeof()` is the flights dataset?

```{r}

```

#### How many observations (rows) are there?

```{r}

```

#### What are the names of the variables?

```{r}

```

#### How many complete cases are there? 

Cases is synonymous with observations (or rows) here.

```{r}

```

#### How many incomplete cases are there? 

```{r}

```


### Building from Scratch

*erase this and add your own notes*

#### build a data.frame

```{r}

```

#### build a tibble

```{r}

```

#### build a list

```{r}

```


***
```{r}
sessionInfo()
```

