---
title: "Worksheet: Introduction to Vectors"
subtitle: "STT 2860 Intro to Data Management & Visualization" 
author: "Jill E. Thomley"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: true
    toc_depth: 4 
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

***

* [Dr. Thomley's Introduction to Vectors slides](https://stat-jet-asu.github.io/Slides/STT2860/IntroVectors.html)
* [Dr. Thomley's Patterned Vectors slides](https://stat-jet-asu.github.io/Slides/STT2860/PatternedVectors.html)

***

### Packages Used

```{r packagesused}
# load all packages used here

```


### Using the `c()` Function

*erase this and add your own notes*

#### logical vector

```{r}

```

#### numeric vector

```{r}

```

#### character vector

```{r}

```

#### mixed vectors

```{r}

```


### Using the `:` Function

*erase this and add your own notes*

#### increasing sequence

```{r}

```

#### decreasing sequence

```{r}

```

#### non-integer sequence

```{r}

```


### Using the `seq()` Function

*erase this and add your own notes*

#### function syntax

    seq(from = 1, to = 1, by = ((to - from)/(length.out - 1)), length.out = NULL, along.with = NULL, ...)

#### using parameter `by =`

```{r}

```

#### using parameter `length.out =`

```{r}

```

#### using parameter `along.with =`

```{r}

```


### Using the `rep`()` Function

#### Create a Sample Vector

Make a vector to use in my examples.

```{r}
text <- c("BOOM", "Shaka", "Laka")
```

#### Repeat using `times =`

```{r}
# rep(c("BOOM", "Shaka", "Laka"), times = 2)
rep(text, times = 2)

repeat1 <- rep(text, times = 2)
```

#### Repeat using `each =`

```{r}
rep(text, each = 2)
```

#### Using both `times =` and `each =`

The function will to each first then times, no matter which comes first in the function.

```{r}
rep(text, times = 3, each = 2)
rep(text, each = 2, times = 3)
```

#### Other Options

We can repeat a non-constant number of times. notice also that we use the `c()` vector to provide input to another function. Here, the c() vector needs to contain as many numbers as the `text` vector has values.

```{r}
rep(text, times = c(1, 2, 5))
```



***
```{r}
sessionInfo()
```
