---
title: "Intro to Reshaping with Separate/Unite and Pivot"
subtitle: "STT2860 Intro to Data Management and Vis S22"
author: "Author: YOUR NAME HERE"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output: 
  html_document: 
    theme: cosmo
    highlight: textmate
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```
<hr>

This worksheet is meant to be used with:

* [RDS CH12 Tidy data](https://r4ds.had.co.nz/tidy-data.html)
* [dplyr: separate and unite](https://youtu.be/ztcsf7G4S78) (video)
* [How to reshape data with tidyr’s new pivot functions](https://youtu.be/gnUgSkKEW5c) (video)
* [Function reference for tidyr](https://tidyr.tidyverse.org/reference/index.html)

## Packages Used

```{r}
library(tidyverse)
library(skimr) # for skim() function
```

## Datasets Used

Example datasets used in **RDS CH12** from the **tidyr** package.

```{r}
table1
table2
table3
table4a
table4b
```

Example dataset from Dr. Thomley's Class Data Files (linked from AsULearn).

```{r}
powerball <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/powerballnumbers.csv")
head(powerball)
```

Example dataset originally from the Kaggle repo [Netflix Movies and TV Shows](https://www.kaggle.com/shivamb/netflix-shows).

```{r}
netflix <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Kaggle/netflix_2021_10_01.csv")
head(netflix)
```

## Separating and Uniting Columns

### `separate()`

I am separating the character `drawdata` variable into three character vectors for `month`, `day`, and `year.` This will also end up overwriting the original year variable, since one of my `into =` columns is `year`.

```{r}
powerball2 <- powerball %>% 
  separate(
    col = drawdate,
    into = c("month", "day", "year"),
    sep = "/"
  )
head(powerball2)
```

```{r}
powerball3 <- powerball %>% 
  separate(
    col = drawdate,
    into = c("month", "day", "year"),
    sep = "/",
    convert = TRUE,
  )
head(powerball3)
```

### `unite()`

What if we got a dataset with month, day, and year separate, but we want them together so we can create a date-time variable.

```{r}
powerball_united <- powerball3 %>% 
  unite(
    col = "datedrawn", 
    month:year,
    sep = "-"
    ) # since month, day, year are together I can use month:year
head(powerball_united)
```

```{r}
powerball_united2 <- powerball3 %>% 
  unite(
    col = "datedrawn", 
    c(year, month, day),    # default R format of year-month-day
    sep = "-"
    ) # since month, day, year are together I can use month:year
head(powerball_united2)
```

## Pivoting the Data "Rectangle"

### `pivot_longer()`

Using `table4a`, I want to make `year` a variable I can work with in R. Right now, the years are variable names (and bad ones, because they should not start with numbers). My final dataset should have three variables: `country`, `year`, and the values from the `1999` and `2000` columns.

Afghanistan  1999   745
Afghanistan  2000  2666
etc...

```{r}
table4a
```

Now `year` will be a piece of information (variable), not variable names.

The old (outdated) function for this was `gather()`. It still works, but has been replaced by `pivot_longer()`.

```{r}
table4a
```

```{r}
table4a_long <- table4a %>% 
  pivot_longer(
    cols = `1999`:`2000`,  # what columns are getting stacked up?
    names_to = "year",     # where will the column names go to?
    values_to = "cases"    # what is the new name of the stacked data
  )
table4a_long
```

```{r}
table4b
```

```{r}
table4b_long <- table4b %>% 
  pivot_longer(
    cols = `1999`:`2000`,       # what columns are getting stacked up?
    names_to = "year",          # where will the column names go to? they represent years
    values_to = "population"    # the values in the stacked columns represent population
  )
table4b_long
```

Now join the two tables!

What are some of the join functions?

* `inner_join()`
* `left_join()`
* `right_join()`
* `full_join()`
* `semi_join()`
* `anti_join()`

```{r}
full_join(
  x = table4a_long, 
  y = table4b_long,
  by = c("country", "year")
  )
```

### `pivot_wider()`

```{r}
table2
```

I need to get cases and opopulation into their own columns so I can calculate `rate = cases / population` using `mutate()`.

```{r}
table2 %>% 
  pivot_wider(
    names_from = type,    # my new columns will be names cases and population
    values_from = count   # what column values are going to fill my new columns
  ) %>% 
  mutate(rate = cases / population * 10000)
```

## Netflix Example

### `separate()` movie categories

Let's first try four categories...

We'll look at a few rows with `head()` and use `skim()` to find `NA` values. Keep the separated column for now so we can verify our code.

```{r}
netflix_4cat <- netflix %>%
  separate(
    col = listed_in,
    into = c("cat1", "cat2", "cat3", "cat4"),
    sep = ",",
    remove = FALSE
  )
```

```{r}
head(netflix_4cat, n = 10)

netflix_4cat %>% 
  select(cat1:cat4) %>% 
  skim()
```

We can reduce the number of categories, since it seems as if no show has a Category 4.

```{r}
netflix_3cat <- netflix %>%
  separate(
    col = listed_in,
    into = c("cat1", "cat2", "cat3"),
    sep = ",",
    remove = FALSE
  )
```

```{r}
head(netflix_3cat, n = 10)

netflix_3cat %>% 
  select(cat1:cat3) %>% 
  skim()
```

Use `pivot_longer()` to put all the categories in the same column.

```{r}
netflix_3cat_long <- netflix_3cat %>% 
  pivot_longer(
    cols = c("cat1", "cat2", "cat3"),
    names_to = "cat_order",
    values_to = "cat_name"
  )
netflix_3cat_long
```

```{r}
netflix_3cat_long %>% 
  count(cat_name) %>% 
  arrange(desc(n))
```

Remove any rows that have `NA` values in the new column you created.

Filter to get rid of rows that have NA as a `cat_name`.

```{r}

```

<hr>
```{r}
sessionInfo()
```
