---
title: "Intro to Binding and Joining in R"
subtitle: "STT2860 Intro to Data Management and Vis S22"
author: "Author: YOUR NAME HERE"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output: 
  html_document: 
    theme: cosmo
    highlight: textmate
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```
<hr>

This worksheet is meant to be used with:

* [RDS CH13 Relational data](https://r4ds.had.co.nz/relational-data.html)
* [Mutating joins](https://dplyr.tidyverse.org/reference/mutate-joins.html)
* [Filtering joins](https://dplyr.tidyverse.org/reference/filter-joins.html)
* [Dr. Thomley's Instructor Datasets](https://github.com/STAT-JET-ASU/Datasets/tree/master/Instructor)

## Packages Used

```{r packagesused}
library(tidyverse)
library(skimr)
```

## Datasets Used

### Myers-Briggs Personality Datasets

Personality types determined using the [Jung Typology Test](https://www.humanmetrics.com/personality).

#### STT1810 Classes

```{r readclass1}
class1 <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/personalitySTT1810.csv")
glimpse(class)
```

#### STT3850 Classes

```{r readclass2}
class2 <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/personalitySTT3850.csv")
glimpse(class)
```

### Census Population Data

Census data taken from tables found on Wikipedia.

#### 1970 and 1980

```{r readcensus1}
census1 <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/census19701980.csv")
glimpse(census1)
```

#### 1990 and 2000

```{r readcensus2}
census2 <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/census19902000.csv")
glimpse(census2)
```

#### 2010 and 2020

```{r readcensus3}
census3 <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/census20102020.csv")
glimpse(census3)
```

### Electoral Votes

Data from [US Electoral College Votes Per State 1788-2020 | Kaggle](https://www.kaggle.com/datasets/daithibhard/us-electoral-college-votes-per-state-17882020)

```{r readelectoralvotes}
votes <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/electoralvotes.csv")
glimpse(votes)
```

## Examples of Combining

### `bind_rows()`

#### Simple row binding

```{r}

```

#### Binding with an ID #1

```{r}

```

#### Binding with an ID #2

```{r}

```

### `bind_cols()`

#### Simple column binding

```{r}

```

#### Duplicate columns?

```{r}

```

### `left_join()`

```{r}

```

### `right_join()`

```{r}

```

### `semi_join()`

```{r}

```

### `anti_join()`

```{r}

```

<hr>
```{r}
sessionInfo()
```
