---
title: "Color/Fill/Faceting/Saving"
author: "YOUR NAME HERE"
date: "Last rendered on `r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: true
    toc_depth: 4
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

### Packages Used

```{r}
library(tidyverse)
library(ggthemes)
library(naniar)
```

### Data Used

The code includes a preview of some skills you will learn later in the course.

#### NFL Player Data

Description: [NFL Players 2016 Dataset](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/NFLplayers2016.html)

```{r}
NFL <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/NFLplayers2016.csv") %>%
  mutate(Playtime = factor(Playtime, levels = c(
    "Rookie",
    "1-4 Yrs",
    "5-8 Yrs",
    "9-12 Yrs",
    "13+ Yrs"
  )))
```

```{r}
glimpse(NFL)
```

```{r}
head(NFL)
```

#### Mauna Loa Data

Source: [NOAA Trends in Atmospheric Carbon Dioxide](https://gml.noaa.gov/ccgg/trends/)

```{r}
MLCO2 <- read_table(
  file = "https://gml.noaa.gov/webdata/ccgg/trends/co2/co2_mm_mlo.txt",
  comment = "#",
  col_names = c(
    "year",
    "month",
    "dec_date",
    "co2",
    "co2_adj",
    "num_days",
    "sd_days",
    "uncertainty"
  )
) %>%
  replace_with_na(replace = list( 
    num_days = -1,
    sd_days = -9.99,
    uncertainty = -0.99
  )) %>%
  filter(year >= 2018)

# replace_with_na() function comes from the naniar package
# https://cran.r-project.org/web/packages/naniar/vignettes/getting-started-w-naniar.html
```

```{r}
glimpse(MLCO2)
```

```{r}
head(MLCO2)
```


### Facets

#### Facet Wrap

Basic Histogram

```{r}
ggplot(NFL, aes(x = Weight)) +
  geom_histogram(bins = 40, color = "black", fill = "orange")
```

Basic Histogram with Fill

```{r}
ggplot(NFL, aes(x = Weight, fill = PositionGrp)) +
  geom_histogram(bins = 40, color = "black")
```

```{r}
ggplot(NFL, aes(x = Weight, fill = Side)) +
  geom_histogram(bins = 40, color = "black")
```

Basic Histogram with Fill and Formatting

https://r-graph-gallery.com/histogram_several_group.html

```{r}
ggplot(NFL, aes(x = Weight, fill = PositionGrp)) +
  geom_histogram(bins = 40, position = "identity", color = "black", alpha = 0.25)
```

```{r}
ggplot(NFL, aes(x = Weight, fill = Side)) +
  geom_histogram(bins = 40, position = "identity", color = "black", alpha = 0.25)
```

Basic Histogram Using Faceting vs. Fill

```{r}
ggplot(NFL, aes(x = Weight)) +
  geom_histogram(bins = 20, color = "black", fill = "orange") +
  facet_wrap(. ~ PositionGrp)
```

```{r}
ggplot(NFL, aes(x = Weight)) +
  geom_histogram(bins = 20, color = "black", fill = "orange") +
  facet_wrap(. ~ Side)
```

Basic Histogram Using Faceting and Fill

```{r}
ggplot(NFL, aes(x = Weight, fill = PositionGrp)) +
  geom_histogram(bins = 20, color = "black") +
  facet_wrap(. ~ PositionGrp)
```

```{r}
ggplot(NFL, aes(x = Weight, fill = Side)) +
  geom_histogram(bins = 20, color = "black") +
  facet_wrap(. ~ Side)
```

Eliminating the Unnecessary Fill Legend

```{r}
ggplot(NFL, aes(x = Weight, fill = PositionGrp)) +
  geom_histogram(bins = 20, color = "black") +
  facet_wrap(. ~ PositionGrp) +
  theme(legend.position = "none")
```

```{r}
ggplot(NFL, aes(x = Weight, fill = Side)) +
  geom_histogram(bins = 20, color = "black") +
  facet_wrap(. ~ Side) +
  theme(legend.position = "none")
```

Setting the Number of Columns for Wrap

```{r}
ggplot(NFL, aes(x = Weight, fill = PositionGrp)) +
  geom_histogram(bins = 20, color = "black") +
  facet_wrap(. ~ PositionGrp, ncol = 2) +
  theme(legend.position = "none")
```

Setting Number of Rows, Different Fill

```{r}
ggplot(NFL, aes(x = Weight, fill = Side)) +
  geom_histogram(bins = 20, color = "black") +
  facet_wrap(. ~ PositionGrp, nrow = 2)
```


#### Facet Grid

    syntax: facet_grid(rowvar ~ colvar)

Basic Histogram Using Faceting and Fill (Columns)

```{r}
ggplot(NFL, aes(x = Weight, fill = Side)) +
  geom_histogram(bins = 20, color = "black") +
  facet_grid(. ~ Side) +
  theme(legend.position = "none")
```

Basic Histogram Using Faceting and Fill (Rows)

```{r}
ggplot(NFL, aes(x = Weight, fill = Side)) +
  geom_histogram(bins = 20, color = "black") +
  facet_grid(Side ~ .) +
  theme(legend.position = "none")
```

Basic Histogram Using Faceting and Fill (Rows and Columns)

```{r}
ggplot(NFL, aes(x = Weight, fill = Side)) +
  geom_boxplot(color = "black") +
  facet_grid(Playtime ~ Side) +
  theme(legend.position = "none")
```


### Multiple Aesthetics/Geometries

https://r-graphics.org/chapter-line-graph
https://r-graphics.org/chapter-scatter
https://rkabacoff.github.io/datavis/Time.html

Using Two Geometries with the Same Aesthetics

```{r}
ggplot(MLCO2, aes(x = dec_date, y = co2)) +
  geom_line() +
  geom_point()
```

Formatting the Two Geometries (Some Options)

https://r-graph-gallery.com/273-custom-your-scatterplot-ggplot2.html
https://blog.albertkuo.me/post/point-shape-options-in-ggplot/
http://www.cookbook-r.com/Graphs/Shapes_and_line_types/

```{r}
ggplot(MLCO2, aes(x = dec_date, y = co2)) +
  geom_line(color = "red", linetype = 2) +
  geom_point(color = "blue", size = 2, shape = 8)
```

Using Two Geometries with Different Aesthetics

```{r}
ggplot(MLCO2, aes(x = dec_date)) +
  geom_line(aes(y = co2), color = "red") +
  geom_point(aes(y = co2), color = "red") +
  geom_line(aes(y = co2_adj)) +
  geom_point(aes(y = co2_adj), size = 0.75)
```

Creating a Theme Object

```{r}
NOAATHEME <- theme_base(base_size = 15) +
  theme(
    plot.title = element_text(
      size = 14,
      family = "Calibri",
      hjust = 0.5,
      margin = unit(c(0, 0, 0, 0), "cm")
    ),
    axis.title = element_text(
      size = 12,
      family = "Calibri"
    ),
    axis.ticks.length = unit(-0.15, "cm"),
    axis.text = element_text(
      size = 10,
      family = "Calibri",
      margin = unit(c(0.4, 0.4, 0.4, 0.4), "cm")
    ),
    plot.background = element_blank()
  )
```

Using the Theme Object in a Plot

```{r}
ggplot(MLCO2, aes(x = dec_date)) +
  geom_line(aes(y = co2), color = "red") +
  geom_point(aes(y = co2), color = "red") +
  geom_line(aes(y = co2_adj)) +
  geom_point(aes(y = co2_adj), size = 0.75) +
  scale_y_continuous(breaks = seq(406, 420, 2)) +
  labs(
    title = expression("Recent Monthly Atmospheric CO"[2] * " at Mauna Loa Observatory"),
    x = "Year",
    y = expression("CO"[2] * " mole fraction (ppm)")
  ) +
  NOAATHEME
```

Using Inventive Formatting to Replicate a Plot
Also Saving the Plot as an Object in RStudio

```{r}
NOAAplot <- ggplot(MLCO2, aes(x = dec_date)) +
  geom_line(aes(y = co2), color = "red") +
  geom_point(aes(y = co2), color = "red") +
  geom_line(aes(y = co2_adj)) +
  geom_point(aes(y = co2_adj), size = 0.75) +
  scale_x_continuous(
    breaks = seq(2018, 2022.75, 0.25),
    labels = c(
      "2018", rep("", 3),
      "2019", rep("", 3),
      "2020", rep("", 3),
      "2021", rep("", 3),
      "2022", rep("", 3)
    ),
    limits = c(2018, 2022.75),
    sec.axis = dup_axis(labels = NULL, name = "")
  ) +
  scale_y_continuous(
    breaks = seq(405.5, 421.5, 0.5),
    labels = c(
      "",
      "406", rep("", 3),
      "408", rep("", 3),
      "410", rep("", 3),
      "412", rep("", 3),
      "414", rep("", 3),
      "416", rep("", 3),
      "418", rep("", 3),
      "420", rep("", 3)
    ),
    limits = c(405.5, 421.5),
    sec.axis = dup_axis(labels = NULL, name = "")
  ) +
  labs(
    title = expression("Recent Monthly Atmospheric CO"[2] * " at Mauna Loa Observatory"),
    x = "Year",
    y = expression("CO"[2] * " mple fraction (ppm)")
  ) +
  NOAATHEME
```

```{r}
NOAAplot
```

Writing a Plot to a File

https://r-coder.com/save-plot-r/

```{r}
# open the png device
png("NOAAPlot.png")

# "create" the plot
NOAAplot

# turn the device off
dev.off()
```

```{r}
pdf("NOAAplotA.pdf", # file name
  width = 7, height = 6, # width and height in inches
  bg = "white", # background color
  colormodel = "cmyk", # color model (cmyk is required for most publications)
  paper = "USr"
) # paper size

ggplot(MLCO2, aes(x = dec_date, y = co2)) +
  geom_line(color = "red", linetype = 2) +
  geom_point(color = "blue", size = 2, shape = 8) +
  theme_classic()

dev.off()
```

Not a Great Plot, but Very Different Geometries

```{r}
ggplot(MLCO2, aes(x = dec_date)) +
  geom_line(aes(y = co2), color = "red") +
  geom_point(aes(y = co2), color = "red") +
  geom_col(aes(y = co2_adj), width = 0.02)
```


<hr>
```{r}
sessionInfo()
```
