---
title: "Add-On Doc"
output: html_document
---

-------------

NEW SECTION 1

-------------

#### Spread the Causes of Death

Spread the data so that the `births` and `deaths` for the two clinics are in different columns.

```{r}

```

#### Vietnam Draft Data

Read the Data
```{r}
draftdata <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/vietnamdraft.csv")
```

Explore the Structure
```{r}
glimpse(draftdata)
head(draftdata)
```

#### Spread Draft Numbers 

Spread the data so that the `draftnumbers` for each `draftyear` are in different columns.

```{r}

```

-------------

NEW SECTION 2

-------------

### Using `separate_wider_*()` and `unite()`

Reference: 

* https://tidyr.tidyverse.org/reference/separate_wider_delim.html
* https://tidyr.tidyverse.org/reference/unite.html

#### Our Class Data

Read the Data
```{r}
classdata <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/STT2860DataScience1/main/Worksheets/stt2860fall2023.csv")
```

Explore the Structure
```{r}
glimpse(classdata)
head(classdata)
```

#### first operation here 

```{r}

```

#### second operation here 

```{r}

```


