---
title: "Worksheet: Introduction to ggplot2 #2"
subtitle: "STT 2860 Intro to Data Mgmt & Visualization" 
author: "PUT YOUR NAME HERE"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

***

### Packages Used

```{r}
# load all packages used here
library(tidyverse)
library(nycflights13)
library(skimr)
```

### The `flights` dataset from nycflights13

This dataset is in the nycflights13 package, so we have access once we load the package.

#### Load the dataset

```{r}
flights <- flights
```

#### Explore the structure with `skim()`

```{r}
skim(flights)
```

### NFL PLayers 2016 Dataset

Description: https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/NFLplayers2016.html

#### Load the dataset

```{r}
NFL <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/NFLplayers2016.csv")
```

#### Explore the structure with `skim()`

```{r}
skim(NFL)
```

### Barplots with `geom_bar()`

For `geom_bar()`, we input an `x` variable and the geometry counts how many of each value occurs in that variable, and that become the "y" value for the height of the bars.

```{r}
ggplot(flights, aes(x = origin)) +
  geom_bar()
```

```{r}
ggplot(flights, aes(x = origin)) +
  geom_bar(color = "black", fill = "orange", width = 0.5) +
  labs(
    title = "Origins of Flights from the New York Metro Area in 2013",
    x = "originating airport by 3-letter code"
  ) +
  theme_minimal()
```

The default order for bars in a plot is alphabetical by category name. Also, sometimes long category names overlap.

```{r}
ggplot(NFL, aes(x = PositionGrp)) +
  geom_bar()
```

Options for fixing category names include:

* change them to make them shorter

    * make it a factor var to change the names
    * use `scale_x_discrete()` to change labels
    
* rotate or slant the text

    * typically with a `theme()` element
    * also with the use `scale_x_discrete()` (I believe)

* flip the plot sideways

    * map `PositionGrp` to the y aesthetic
    * we could use `coord_flip()` to flip `x` and `y`

```{r}
ggplot(NFL, aes(y = PositionGrp)) +
  geom_bar() +
  scale_x_continuous(breaks = seq(from = 0, to = 800, by = 100))
```

```{r}
ggplot(NFL, aes(x = PositionGrp)) +
  geom_bar() +
  scale_y_continuous(breaks = seq(from = 0, to = 800, by = 100)) +
  coord_flip()
```

Where do flights from the three airports go?

```{r}
ggplot(flights, aes(x = origin, fill = dest)) +
  geom_bar()
```

```{r}
ggplot(NFL, aes(x = PositionGrp, fill = Side)) +
  geom_bar() +
  coord_flip()
```

```{r}
ggplot(NFL, aes(x = Side, fill = PositionGrp)) +
  geom_bar(position = "dodge", color = "black")
```

```{r}
ggplot(NFL, aes(x = Side, fill = PositionGrp)) +
  geom_bar(position = "fill", color = "black") +
  labs(y = "proportion")
```

### Barplots with `geom_col()`

For `geom_col()`, we input an `x` variable and we input a `y` variable that specifies the heights of the bars. The data is already summarized somehow.

I can count the categories up myself to generate a smaller dataset with categories and counts. I get a dataset with two variables: the variable I counted, and one called `n`
that is the counts.

```{r}
NFL_PosGrp_Counts <- NFL %>% 
  count(PositionGrp)

NFL_PosGrp_Counts
```

```{r}
ggplot(NFL_PosGrp_Counts, aes(x = PositionGrp, y = n)) +
  geom_col()
```

If I make the plot this way, I can rearrange the bars in order using the `reorder()` function. Here, I want to change the order of the categories according to their counts.

```{r}
ggplot(NFL_PosGrp_Counts, aes(x = reorder(PositionGrp, n), y = n)) +
  geom_col() 
```

```{r}
ggplot(NFL_PosGrp_Counts, aes(x = reorder(PositionGrp, n), y = n)) +
  geom_col() +
  coord_flip()
```

```{r}
ggplot(NFL_PosGrp_Counts, aes(x = reorder(PositionGrp, -n), y = n)) +
  geom_col() 
```

```{r}
ggplot(NFL_PosGrp_Counts, aes(x = reorder(PositionGrp, -n), y = n)) +
  geom_col() +
  coord_flip()
```

```{r}
NFL_wts <- NFL %>% 
  group_by(Side) %>% 
  summarize(mean_wt = mean(Weight))

NFL_wts

ggplot(NFL_wts, aes(x = Side, y = mean_wt)) +
  geom_col(width = 0.5)
```

### The `annotate()` function



***
```{r}
sessionInfo()
```

