---
title: "Worksheet: Introduction to Plotting with ggplot2"
subtitle: "STT 2860 Intro to Data Mgmt & Visualization" 
author: "PUT YOUR NAME HERE"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: true
    toc_depth: 4 
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

***

### Packages Used

```{r}
# load all packages used here
library(tidyverse)
```


### Datasets Used

#### Popular Kids

The dataset `popular.csv` is from [Dr. Thomley's Data Files](https://stat-jet-asu.github.io/Datasets/DatasetList.html). You can read about it [HERE](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/popular.html).

```{r}
# import the dataset from the web URL
kids <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/popular.csv")
```

```{r}
# explore the dataset
```

#### Semmelweis Data

The dataset `semmelweisannualbyclinic.csv` is from [Dr. Thomley's Data Files](https://stat-jet-asu.github.io/Datasets/DatasetList.html). You can read about Dr. Ignaz Semmelweis [HERE](https://www.pbs.org/newshour/health/ignaz-semmelweis-doctor-prescribed-hand-washing). This dataset contains annual observations for two different clinics that led him to implement his handwashing experiment.

```{r}
semm <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/semmelweisannualbyclinic.csv")
```

```{r}
# explore the dataset
```

#### Flights Data

The `flights` dataset is from the nycflights13 package, which is installed on the RStudio server.

```{r}
# save the dataset to your Environment
```

```{r}
# explore the dataset
```


### Using `geom_bar()`

#### Basic Bar Plot

```{r}

```


### Using `geom_col()`

#### Subheader Here!



***
```{r}
sessionInfo()
```

