---
title: "Worksheet: Introduction to ggplot2"
subtitle: "STT 2860 Intro to Data Mgmt & Visualization" 
author: "PUT YOUR NAME HERE"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: true
    toc_depth: 4 
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

***

Dr. Thomley's Data Files: https://stat-jet-asu.github.io/Datasets/DatasetList.html

The tidyverse ggplot2 page: https://ggplot2.tidyverse.org/

***

### Packages Used

```{r}
# load all packages used here

```


### Anthropometric Dataset

Description: https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/anthropometric.html 

#### Load the dataset

```{r}

```

#### Explore the structure

```{r}

```


### Atlantic Hurricanes Dataset

Description: https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/atlantichurricanes18512023.html

#### Load the dataset

```{r}

```

#### Explore the structure

```{r}

```


### NFL PLayers 2016 Dataset

Description: https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/NFLplayers2016.html

#### Load the dataset

```{r}

```

#### Explore the structure

```{r}

```


### Scatterplots

*add some notes here*

#### Minimal Scatterplot

```{r}

```

#### Additional Aesthetics

color

```{r}

```

shape

```{r}

```

size

```{r}

```

#### Non-Aesthetic Formatting

```{r}

```

#### Labeling Your Plots!

```{r}

```

WE WILL ADD MORE HERE AS WE GO ALONG!


***
```{r}
sessionInfo()
```

