---
title: "Worksheet: Introduction to Vectors"
subtitle: "STT 2860 Intro to Data Management & Visualization" 
author: "PUT YOUR NAME HERE"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: true
    toc_depth: 4 
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

***

* [Dr. Thomley's Introduction to Vectors slides](https://stat-jet-asu.github.io/Slides/STT2860/IntroVectors.html)
* [Dr. Thomley's Patterned Vectors slides](https://stat-jet-asu.github.io/Slides/STT2860/PatternedVectors.html)

***

### Packages Used

```{r packagesused}
# load all packages used here

```


### Using the `c()` Function

*erase this and add your own notes*

#### logical vector

```{r}

```

#### numeric vector

```{r}

```

#### character vector

```{r}

```

#### mixed vectors

```{r}

```


### Using the `:` Function

*erase this and add your own notes*

#### increasing sequence

```{r}

```

#### decreasing sequence

```{r}

```

#### non-integer sequence

```{r}

```


### Using the `seq()` Function

*erase this and add your own notes*

#### function syntax

    seq(from = 1, to = 1, by = ((to - from)/(length.out - 1)), length.out = NULL, along.with = NULL, ...)

#### using parameter `by =`

```{r}

```

#### using parameter `length.out =`

```{r}

```

#### using parameter `along.with =`

```{r}

```


### Using the `rep`()` Function

*erase this and add your own notes*

*erase this and add your own subsections*


***
```{r}
sessionInfo()
```

