---
title: "Worksheet: Pivoting!"
subtitle: "STT 2860 Intro to Data Mgmt & Visualization" 
author: "PUT YOUR NAME HERE"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: true
    toc_depth: 4 
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

***

### Packages Used

```{r}
library(tidyverse)
library(nycflights13)
library(fivethirtyeight)
```

### Democracy Score Data

```{r}
dem_score <- read_csv("https://moderndive.com/data/dem_score.csv")
dem_score
```

### Drinks & Drinks Smaller Data

```{r}
drinks <- drinks
drinks
```

```{r}
drinks_smaller <- drinks %>% 
  filter(country %in% c("USA", "China", "Italy", "Saudi Arabia")) %>% 
  select(-total_litres_of_pure_alcohol) %>% 
  rename(beer = beer_servings, spirit = spirit_servings, wine = wine_servings)
drinks_smaller
```

### Airline Safety Dataset

```{r}
airline_safety <- airline_safety
glimpse(airline_safety)
```

### Pivoting to a Longer Form

#### Smaller Drinks Dataset

```{r}
drinks_smaller_tidy <- pivot_longer(
  drinks_smaller,
  cols = beer:wine,
  names_to = "type",
  values_to = "servings"
)
drinks_smaller_tidy
```

#### Full Drinks Dataset

```{r}
drinks_tidy <- drinks %>% 
  rename(beer = beer_servings, spirit = spirit_servings, wine = wine_servings) %>% 
  select(-total_litres_of_pure_alcohol) %>%   
  pivot_longer(
    cols = beer:wine,
    names_to = "type",
    values_to = "servings"
)
glimpse(drinks_tidy)
```

#### Democracy Score Dataset

```{r}
# dem_score %>% 
#   pivot_longer(
#     cols = `1952`:`1992`,
#     names_to = "year",
#     values_to = "score"
#   )

dem_score_tidy <- dem_score %>% 
  pivot_longer(
    cols = !country,
    names_to = "year",
    values_to = "score"
  ) %>% 
  mutate(year = as.numeric(year))

glimpse(dem_score_tidy)
```

I would like to change as little code as possible when switching from country to country.

```{r}
# input choice of country here
# then I don't need to change
# the name in any code below
countrychoice <- "Russia"

# filter country data to plot
# my dataset is names country
# it contains only the country I want
country <- dem_score_tidy %>%
  filter(country == countrychoice)

# pull out the vector of years
# we will use this in the plot
year_vec <- country %>%
  pull(year)

# pull out the vector of scores
# we will use this in the plot
score_vec <- country %>%
  pull(score)

# create fill vector for labels
# different colors for pos/neg
# we will use this in the plot
score_fill_vec <- country %>%
  mutate(score_fill = ifelse(score < 0, "pink", "lightgreen")) %>%
  pull(score_fill)

# pull out chosen country name
# we will use this in the plot
countryname <- country %>% 
  distinct(country) %>% 
  pull(country)

# create the plot from inputs above!
country %>%
  ggplot(aes(x = year, y = score)) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    size = 0.25
  ) +
  geom_line() +
  scale_x_continuous(
    breaks = seq(1950, 1995, 5),
    limits = c(1949.5, 1997),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    breaks = seq(-10, 10, 2),
    limits = c(-12.5, 12.5)
  ) +
  annotate(
    "label",
    x = year_vec,
    y = score_vec,
    label = score_vec,
    fill = score_fill_vec,
  ) +
  annotate(
    "label",
    x = 1950,
    y = c(-12, +12),
    hjust = 0,
    label = c("highly autocratic", "highly democratic"),
    fill = c("pink", "lightgreen"),
    alpha = 0.25
  ) +
  labs(
    title = paste("Changes in Democracy Score Over Time for", countryname),
    y = "democracy score"
  ) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank())
```

#### Airline Datset Manips

```{r}
airline_safety_smaller <- airline_safety %>% 
  select(airline, starts_with("fatalities"))
airline_safety_smaller
```

```{r}
airline_safety_smaller %>% 
  pivot_longer(
    cols = starts_with("fatalities"),
    names_to = "time_period",
    names_prefix = "fatalities_",
    values_to = "num_fatalities",
    
  )
```

### Watauga County Weather Days

```{r}
WCS <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/wcsweatherdays.csv") %>% 
  select(-total_days, -fall)

glimpse(WCS)
```

```{r}
head(WCS)
```

```{r}
WCS %>% 
  separate_wider_delim(
    year,
    delim = "-",
    names = c("start_year", "end_year"),
    cols_remove = FALSE
  ) 
```

```{r}
WCS2 <- WCS %>% 
  separate_wider_delim(
    year,
    delim = "-",
    names = c("start_year", "end_year")
  ) %>% 
  pivot_longer(
    cols = ends_with("days"),
    names_to = "month",
    values_to = "num_days"
  )
```

```{r}
WCS2 %>% 
  unite(
    "academic_year", 
    start_year:end_year, 
    sep = " to ",
    remove = FALSE)
```

### Skyscrapers Data

```{r}
sky <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/USskyscrapers.csv")

glimpse(sky)
head(sky)
```

```{r}
library(skimr)
skim(sky)
```

```{r}
sky %>% 
  count(Use)
```

```{r}
sky2 <- sky %>% 
  separate_wider_delim(
    Use,
    delim = " / ",
    names = c("Use01", "Use02", "Use03", "Use04"),
    too_few = "align_start",
    cols_remove = TRUE
  ) 
```

```{r}
glimpse(sky2)
```

```{r}
sky2 %>% 
  count(Use04)
```

```{r}
sky3 <- sky2 %>% 
  pivot_longer(
    cols = Use01:Use04,
    names_to = "UseNum",
    values_to = "Use"
  ) %>% 
  filter(!is.na(Use)) %>% 
  select(-UseNum)
```

```{r}
sky3
```

```{r}
sky %>% 
  separate_longer_delim(
    Use,
    delim = " / "
  ) 
```

### Joining `flights` dataset

Are there any cases in `flights` that do not have a match in `airlines`?

```{r}
flights %>% 
  anti_join(airlines, by = "carrier")
```

```{r}
flights2 <- flights %>% 
  left_join(airlines, by = "carrier") %>% 
  relocate(name, .after = carrier) %>% 
  rename(carrier_name = name)
```

```{r}
flights3 <- flights2 %>%
  left_join(airports, by = c("origin" = "faa"))
```


***
```{r}
sessionInfo()
```

