---
title: "PR01 NFL 2016 Data Visualizations"
subtitle: "STT2860 Spring 2023"
author: "GROUP MEMBER NAMES HERE"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    code_folding: hide
---
<hr>

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

### Packages Used

```{r}

```


### NFL Players 2016 Dataset

Go to the [description](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/NFLplayers2016.html) of Dr. Thomley's NFL Players 2016 Dataset and read the brief description and variable definitions. Click on the `NFLplayers2016.csv` link at the top and use the URL to read the data into `R`. Save the dataset as an object named `nfl2016`.

```{r nfl2016}
dataURL <- "https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/NFLplayers2016.csv"

nfldata <- read_csv(dataURL) %>%
  mutate(Playtime = factor(Playtime, levels = c("Rookie", "1-4 Yrs", "5-8 Yrs", "9-12 Yrs", "13+ Yrs")))
```

Use `glimpse()` to explore the structure of the `nfldata` dataset.

```{r}

```

Use `head()` to explore the structure of the `nfldata` dataset.

```{r}

```


### Plot #1: Bar Plot of Playtime

* Create a bar plot for the variable `Playtime`. 
* Change the bar colors from gray to some other color/colors.
* Modify the x-axis so that `Yrs` is not abbreviated in the labels.
* Scale the y-axis so it has tick marks every 100 units (e.g., 100, 200, etc.).
* Add an informative title and appropriate labels for both axes.
* You may optionally add a subtitle or caption for more clarity.
* Add the built-in `theme_light()` from ggplot2 to your plot.
* Your final plot should *not* have a legend on it.

```{r}

```

Do you think most NFL players should look forward to long careers, based on your plot?

**ANSWER:** type your answer here


### Plot #2: Histogram of Weight

* Create a histogram with 25 bins for the variable `Weight`. 
* Use `fill =` to include information from the variable `Side`. 
* Add formatting in the geometry so the bars are outlined in black.
* Use `scale_fill_manual()` to change the colors used in the bars.
* Alternatively, you can use `scale_fill_brewer()` to change bar colors.
* Add the built-in `theme_minimal()` from ggplot2 to your histogram.
* Use `theme()` to move the legend to the bottom and remove its title.
* Add an informative title and appropriate labels for both axes.
* You may optionally add a subtitle or caption for more clarity.
* Be sure to include any units of measure in the axis labels.

```{r}

```

Discuss two group comparisons and/or interesting features of this plot.

**ANSWER:** type your answer here


### Plot #3: Density Plot of BMI

Body mass index (BMI) is a measure of body fat based on height and weight. The formula for calculating BMI is BMI = kg/m^2^ where kg is weight in kilograms and m^2^ is height in meters squared. The "healthy" range for BMI is 18.5 to 24.9; BMI > 25 is "overweight"; 30-39 is "obese"; and 40+ "morbidly obese". 

* Use `mutate()` to compute BMI for the football players. 
* Save the dataset containing the new variable as `nfldataBMI`.
* Create a density (i.e., KDE) plot for `BMI` with a bandwidth of 0.75. 
* Use `fill =` and `alpha =` in the geometry to add semi-transparent color.
* Add dashed vertical lines in a contrasting color or colors at 25, 30, and 40.
* Scale the x-axis so it has tick marks at every whole unit (e.g., 24, 25, 26, etc.).
* Additionally, use `limits =` so that the x-axis displays values between 18 and 48.
* Use `annotate()` to label the healthy, overweight, obese, and morbidly obese regions.
* Add an informative title and appropriate labels for both axes.
* You may optionally add a subtitle or caption for more clarity.

```{r}

```

Many people suggest that BMI is a poor measure for athletes, who often have a high body weight due to muscle mass versus fat. What does your plot suggest about this?

**ANSWER:** type your answer here


### Plot 4: Violin and Box Plots of Age

* We want to compare `Age` among the three `Side` groups. 
* Use `geom_violin()` as the first geometry layer of the plot.
* Add `geom_boxplot()` as the second geometry layer of the plot.
* Adjust the width of the boxplots so they are completely inside the violins.
* Fill the boxplots with a color to visually contrast with the violins.
* Increase the size of the boxplot outliers for enhanced visibility.
* Add an informative title and appropriate labels for both axes.
* You may optionally add a subtitle or caption for more clarity.
* Be sure to include any units of measure in the axis labels.
* Add any other formatting you feel will enhance meaning.

```{r}

```

Discuss two group comparisons and/or interesting features of this plot.

**ANSWER:** type your answer here


### Plot 5: Scatterplots of Weight vs. Height

#### Offensive Players

* Use `Side` to `filter()` the dataset to choose `Offense` players only. 
* Save this new dataset containing the offensive players as `offdata`. 
* Create a scatterplot of player weight (y) versus player height (x).
* Use the `color =` aesthetic to add information from `PositionGrp`. 
* Add an appropriate title and better labels for the x and y axes.
* Be sure to include any units of measure in the axis labels.

```{r}

```

#### Defensive Players

* Use `Side` to `filter()` the dataset to choose `Defense` players only. 
* Save this new dataset containing the defensive players as `defdata`. 
* Create a scatterplot of player weight (y) versus player height (x).
* Use the `color =` aesthetic to add information from `PositionGrp`. 
* Add an appropriate title and better labels for the x and y axes.
* Be sure to include any units of measure in the axis labels.

```{r}

```

Does the relationship between weight and height seem to differ based on position grouping? Discuss.

**ANSWER:** type your answer here


### Plot 6: Your Choice!

Create one more plot of your choice using the `nfldata`. Explain what you were intending to show/explore with your plot.

```{r}

```

**EXPLANATION:**


<hr>

```{r}
sessionInfo()
```

