---
title: "Example Markdown Document"
subtitle: "STT2860 Intro to Data Management and Vis S22"
author: "Author: Jill E. Thomley"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output: 
  html_document: 
    theme: cosmo
    highlight: textmate
    toc: true
    toc_depth: 4
    code_folding: hide
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```
<hr>

```{r}
library(tidyverse)
library(expss)
```


## Mauna Loa CO2 Data

The National Oceanic & Atmospheric Administration (NOAA) monitors [trends in atmospheric carbon dioxide](https://gml.noaa.gov/ccgg/trends/) and makes its data available to the public. For example, we can look at the [monthly averages](https://gml.noaa.gov/webdata/ccgg/trends/co2/co2_mm_mlo.txt) from the observatory on Mauna Loa in Hawaii. The website offers a downloadable csv file, but we can read the data in directly from the text file on the website using `R`. This enhances reproducibility in our analysis, because NOAA is often updating and/or correcting their values.

### Load & Clean the Data

```{r}
mlco2 <- read_table("https://gml.noaa.gov/webdata/ccgg/trends/co2/co2_mm_mlo.txt",
  comment = "#",
  col_names = c(
    "year",
    "month",
    "dec_date",
    "co2",
    "co2_adj",
    "num_days",
    "sd_days",
    "uncertainty"
  )
) %>%
  na_if(lt(0)) %>%
  pivot_longer(cols = co2:co2_adj, names_to = "co2_type", values_to = "ppm")
glimpse(mlco2)
```

### Plots of Trends In CO2

#### Since 1958

```{r}
mlco2 %>%
  ggplot(aes(x = dec_date, y = ppm, color = co2_type)) +
  geom_line() +
  labs(
    title = expression("Atmospheric Mean CO"[2] * " at Mauna Loa Observatory"),
    caption = "source: https://gml.noaa.gov/ccgg/trends/",
    x = "year",
    y = "parts per million (ppm)",
    color = "CO2 Measurement"
  ) +
  scale_x_continuous(breaks = seq(from = 1960, to = 2020, by = 5)) +
  scale_y_continuous(breaks = seq(from = 320, to = 420, by = 20)) +
  scale_color_manual(
    labels = c("Monthly Avg CO2", "Seasonally Adj CO2"),
    values = c("red", "black")
  ) +
  theme_linedraw()
```

#### Since 2016

```{r}
mlco2 %>%
  filter(year >= 2016) %>% 
  ggplot(aes(x = dec_date, y = ppm, color = co2_type)) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(
    title = expression("Recent Monthly Mean CO"[2] * " at Mauna Loa Observatory"),
    caption = "source: https://gml.noaa.gov/ccgg/trends/",
    x = "year",
    y = "parts per million (ppm)",
    color = "CO2 Measurement"
  ) +
  scale_x_continuous(breaks = 2016:2022) +
  scale_y_continuous(breaks = seq(from = 400, to = 420, by = 5)) +
  scale_color_manual(
    labels = c("Monthly Avg CO2", "Seasonally Adj CO2"),
    values = c("red", "black")
  ) +
  theme_linedraw()
```

Reference:

[Trends in Atmospheric Carbon Dioxide](https://gml.noaa.gov/ccgg/trends/)

<hr>
```{r}
sessionInfo()
```
