---
title: "Intro to Visualization with ggplot2 & Other Packages"
subtitle: "STT2860 Intro to Data Management and Vis S22"
author: "Author: YOUR NAME HERE"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: true
    toc_depth: 4
    toc_float: true
    code_folding: hide
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```
<hr>

## Packages Used

```{r}
library(tidyverse)
# library(treemapify)
library(ggridges)
library(waffle)
```

## Datasets Used

### Capital Punishment

```{r}
cap <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/cappunish20220107.csv")
glimpse(cap)
```

### Mauna Loa CO2 Data

```{r}
mlco2 <- read_table("https://gml.noaa.gov/webdata/ccgg/trends/co2/co2_mm_mlo.txt",
  comment = "#",
  col_names = c(
    "year",
    "month",
    "dec_date",
    "co2",
    "co2_adj",
    "num_days",
    "sd_days",
    "uncertainty"
  )
) %>%
  select(-(num_days:uncertainty)) %>% 
  pivot_longer(cols = co2:co2_adj, names_to = "co2_type", values_to = "ppm")
glimpse(mlco2)
```

### MLB Salaries 1985-2015

```{r}
mlb <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/MLBsalaries19852015.csv")
glimpse(mlb)
```

```{r}
mlb_yr <- mlb %>%
  group_by(Year) %>%
  summarize(avg = mean(Salary),
            med = median(Salary))
glimpse(mlb_yr)
```

```{r}
mlb_yr_long <- mlb_yr %>%
  pivot_longer(avg:med, names_to = "stat", values_to = "dollars")
glimpse(mlb_yr_long)
```

### NFL Players 2016

```{r}
nfl <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/NFLplayers2016.csv") %>%
  mutate(Playtime = factor(Playtime,
    levels = c(
      "Rookie",
      "1-4 Yrs",
      "5-8 Yrs",
      "9-12 Yrs",
      "13+ Yrs"
    )
  ))
glimpse(nfl)
```

## Barplots

### One Categorical Variable

```{r}

```

### Two Categorical Variables

#### Stack (Default)

```{r}

```

#### Dodge (Grouped)

```{r}

```

#### Fill (Segmented)

```{r}

```

### Summary Statistics

```{r}

```

## Quantitative Plots

### Histogram/Polygon

```{r}

```

### Density/Ridgeline

```{r}

```

### Boxplots/Violins

#### Boxplot

```{r}

```

#### Violin

```{r}

```

#### Boxplot & Violin

```{r}

```

## Scatterplots/Lines

### Two Variables

```{r}

```

### Time Series

#### Point

```{r}

```

#### Line

```{r}

```

#### Point & Line

```{r}

```

## Other Plots

### Tree Map

```{r}

```

### Waffle

```{r}

```

### TBA

```{r}

```


<hr>
```{r}
sessionInfo()
```
