---
title: "Introduction to Vectors Worksheet"
subtitle: "STT2860 Intro to Data Management and Vis S22"
author: "Author: YOUR NAME HERE"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output: 
  html_document: 
    theme: cosmo
    highlight: textmate
    toc: true
    toc_depth: 4
    toc_float: true
    code_folding: hide
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```
<hr>

This worksheet is meant to be used with the [Introduction to Vectors](https://stat-jet-asu.github.io/Slides/GeneralR/IntroVectors.html#1) class slides and class lecture.

## Atomic Vectors

*put your own notes about atomic vectors here*

## Creating Vectors

### The `vector()` Function

#### Default mode

    vector(mode = "logical", length = 0)

```{r}

```

#### Changing mode and length

Numeric

```{r}

```

Logical

```{r}

```

Integer

```{r}

```

Double

```{r}

```

Character

```{r}

```

### Related Functions

    logical(length = 0)
    integer(length = 0)
    double(length = 0)
    numeric(length = 0)
    character(length = 0)

```{r}

```

## Patterned Vectors

### The `c()` Function

```{r}

```

### The `:` Function

```{r}

```

### The `seq()` Function

```{r}

```

### The `rep()` Function

```{r}

```

## Attributes

```{r}

```

## Creating Lists

```{r}

```

## Creating Tibbles

```{r}

```

<hr>
```{r}
sessionInfo()
```
