---
title: "NFL Players 2016 Data"
subtitle: "STT2860 Project 01 Fall 2022"
author: "GROUP MEMBER NAMES HERE"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---
<hr>

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

### Packages Used

```{r}

```


### Read the Dataset

Go to the [description](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/NFLplayers2016.html) of Dr. Thomley's NFL Players 2016 Dataset and read the brief description and variable definitions. Click on the `NFLplayers2016.csv` link at the top and use the URL to read the data into `R`. Save the dataset as an object named `nfl2016`.

```{r}

```

Use `glimpse()` to explore the structure of your dataset.

```{r}

```


### Plot #1: Histogram of Weight

Create a histogram for the variable `Weight`. Use the `fill` attribute (similar to the `size` and `shape` attributes you learned about in DataCamp) to include information from the variable `Side`. Add an appropriate title.

```{r}

```

Discuss two group comparisons and/or interesting features of this plot.

**ANSWER:** type your answer here


### Plot #2: Histogram of BMI

Body mass index (BMI) is a measure of body fat based on height and weight. The formula for calculating BMI is BMI = kg/m^2^ where kg is weight in kilograms and m^2^ is height in meters squared. The "healthy" range for BMI is 18.5 to 24.9. BMI > 25 is "overweight", 30-39 is "obese", and 40+ "morbidly obese". Use `mutate()` to compute BMI for the football players. Save the dataset as an object named `nfl2016BMI`.

```{r}

```

Create a histogram for `BMI`. Add an appropriate title.

```{r}

```

Many people suggest that BMI is a poor measure for athletes, who often have a high body weight due to muscle mass versus fat. What does your plot suggest about this?

**ANSWER:** type your answer here


### Plot 3: Boxplots of Age

Create a boxplot to compare `Age` among the three `Side` groups. Add an appropriate title.

```{r}

```

Discuss two group comparisons and/or interesting features of this plot.

**ANSWER:** type your answer here


### Plot 4: Scatterplots of Weight vs. Height

Use `Side` to `filter()` the dataset for offense-side players only. Call this new dataset `off2016`. Create a scatterplot of weight (y) versus height (x) and `color` the points according to `PositionGrp`. Add an appropriate title.

```{r}

```

Use `Side` to `filter()` the dataset for defense-side players only. Call this new dataset `def2016`. Create a scatterplot of weight (y) versus height (x) and `color` the points according to `PositionGrp`. Add an appropriate title.

```{r}

```

Does the relationship between weight and height seem to differ based on position grouping? Discuss.

**ANSWER:** type your answer here


### Save Your Data Files

Save the data files `nfl2016BMI`, `off2016`, and `def2016` into an `.Rdata` file called `nfl2016datafiles.Rdata`.

```{r}

```

<hr>

```{r}
sessionInfo()
```

