---
title: "Mauna Loa CO2 Data"
subtitle: "STT2860 Project 01 Fall 2022"
author: "SOLUTIONS // Jill E. Thomley"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---
<hr>

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

### Packages Used

```{r}
library(tidyverse)
```


### Read Datasets

In class we read in monthly carbon dioxide data from NOAA's Mauna Loa observatory. They actually offer several local and global [datasets for download](https://gml.noaa.gov/ccgg/trends/data.html) on their website. 

Read in the **Mauna Loa CO2 annual mean data** from its online (text) page. Give the dataset a sensible name.

```{r}
CO2mean <- read_table(
  file = "https://gml.noaa.gov/webdata/ccgg/trends/co2/co2_annmean_mlo.txt",
  comment = "#",
  col_names = c(
    "year",
    "mean",
    "unc"
  )
)
```

Use `glimpse()` to explore the structure of your dataset.

```{r}
glimpse(CO2mean)
```

Download the (CSV) file for the **Mauna Loa CO2 annual mean growth rates** and read it in (you will have to upload it to RStudio). Give the dataset a sensible name.

```{r}
CO2grow <- read_csv("co2_gr_mlo.csv",
  skip = 58,
  col_names = c(
    "year",
    "increase",
    "unc"
  )
)
glimpse(CO2grow)
```


### Highest Years

Use `arrange()` and `head()` to display the ten years with the highest mean CO2.

```{r}
CO2mean %>% 
  arrange(desc(mean)) %>% 
  head(n = 10)
```

Use `arrange()` and `tail()` to display the ten years with the largest increases.

```{r}
CO2grow %>% 
  arrange((increase)) %>% 
  tail(n = 10)
```

<hr>

```{r}
sessionInfo()
```

