---
title: "Project 1: Individual Coding & Analysis"
subtitle: "STT 2860 Intro to Data Mgmt & Visualization" 
author: "PUT YOUR NAME HERE"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: true
    toc_depth: 4 
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```
***

### Packages Used

```{r}
library(tidyverse)
```

### Datasets Used

You can read the dataset descriptions and variable definitions [HERE](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/cappunish.html).

#### Capital Punishment Dataset

```{r readcappunish}
cpdata <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/cappunish20231018.csv")
```

#### Size of Death Row Dataset

```{r readdeathrowsize}
drdata <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/deathrowsize.csv")
```

#### Death Sentences in NC Dataset

```{r readsentencesNC}
ncdata <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/deathsentencesNC.csv")
```

### Plot #1: How Many States Have Capital Punishment? 

This plot uses the **Capital Punishment** dataset.

* Create a barplot for the variable `dp2`. 
* Use `color =` and `fill =` to format the bars.
* Use `labs()` to add a title and better axis labels.
* Apply a built-in theme from **ggplot2** or **ggthemes**.
* This plot should *not* have a visible legend anywhere on it.
* You may optionally add other formatting you deem useful.

```{r plot1}

```

### Plot #2: Capital Punishment By Region of the U.S.

This plot uses the **Capital Punishment** dataset.

* Create a barplot for the variable `region`. 
* Use `fill =` to shade each bar according to `dp2`.
* Change the `position =` to set each bar equal to 100%.
* Use `width =` to make the bars narrower than the default.
* Use `labs()` to add a title, axis labels, and a better legend label.
* Use `scale_fill_manual()` to choose a different color scheme for the plot. 
* Use `theme()` to move the legend to the bottom of the plot instead of the right
* You may optionally add a subtitle, caption, or ggplot2 built-in theme to the plot.

```{r plot2}

```

### Plot #3: Number of Prisoners on Death Row

This plot uses the **Size of Death Row** dataset.

* Create a barplot for the total number of prisoners (y) in each year (x).
* Fill the bars using the variable `decade` and outline them all in black.
* Set the width of all the bars so there is no visible gap in between them.
* Scale the x-axis to include a tick mark at each decade (1970, 1980, etc).
* Scale the y axis so that it is numbered 0, 500, 1000, 1500, 2000, etc.
* Use `labs()` to add or change the title, axis labels, and legend label. 
* Use `theme()` to do the following:
    * center your title
    * set the plot background to white or blank
    * set all axis tick marks to transparent or blank
    * set the major x-axis gridlines to transparent or blank
    * set the major y-axis gridlines to light gray
* You may optionally change the color scheme used for the bars in your plot

```{r plot3}

```

### Plot #4: Number of Executions Since 1976

This plot uses the **Capital Punishment** dataset.

* Create a violin plot for the variable `post1976` (use `y = 0`).
* Add a boxplot layer and set `width` to fit inside the violin plot.
* Fill each plot with a color of your choice (be colorblind-aware!).
* Scale the x-axis of the plot so that it counts in increments of 25.
* Remove all y-axis markings/gridlines using `scale_y_continuous()`.
* Use `annotate()` to identify the large outlier using the state name.
* Add a title and x-axis label (make sure we know it's all states *and* DC).
* Add the built-in theme `theme_linedraw()` or `theme_minimal` to the plot.

Hint: You can use the `NULL` vector in functions like `scale_y_continuous()`

```{r plot4}

```

### Plot #5: Death Sentences in NC Since 1976

This plot uses the **Death Sentences in NC** dataset.

Create a time series plot of death sentences in North Carolina. Use three geometries on this plot: `geom_area()`, `geom_line()` and `geom_point()`. Use `geom_vline()` to add a vertical line for 2001. The significance of this year is that the General Assembly modified the death penalty statute in several ways, including giving prosecutors the discretion to *not* seek the death penalty in cases where it previously would have been mandatory. Title, label, and format your plot in such a way that the information it conveys would be clear to the reader (i.e., the plot is reasonably stand-alone).

```{r plot5}

```

### Information Conveyed by Your Plots

List two pieces of information you can learn or comparisons you can make from each plot above.

#### Plot #1

* Item 1:
* Item 2:

#### Plot #2

* Item 1:
* Item 2:

#### Plot #3

* Item 1:
* Item 2:

#### Plot #4

* Item 1:
* Item 2:

#### Plot #5

* Item 1:
* Item 2:


***
```{r}
sessionInfo()
```
