---
title: "Project 01 Individual Component"
subtitle: "STT2860 Spring 2023"
author: "YOUR NAME HERE"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    code_folding: hide
---
<hr>

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

### Packages Used

```{r}

```


### Hurricanes Dataset

#### Explore the Data

You can read description of the dataset [HERE](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/atlantichurricanes18512022.html).

```{r nfl2016}
storms <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/atlantichurricanes18512022.csv")
```

Use `glimpse()` to explore the structure of the `storms` dataset.

```{r}

```

#### Time Series Plot of Storms

* The x-axis of the plot in all cases will be the time variable `year`.
* Plot the number of named storms (y) per year (x) using `geom_line()`.
* Add the trend in the number of named storms (y) using `geom_smooth()`.
* Choose and set a color for both the line and trend of named storms.
* Plot the number of major storms (y) per year (x) using `geom_line()`.
* Add the trend in the number of major storms (y) using `geom_smooth()`.
* Choose and set a color for both the line and trend of major storms.
* Add an informative title and appropriate labels for both axes.
* You may optionally add a subtitle or caption for more clarity.
* Add any other formatting you feel will enhance meaning.

```{r}

```

Do you notice any trend(s) in the number of named storms or major hurricanes? Discuss.

**ANSWER:** type your answer here


### Unemployment Dataset

#### Explore the Data

You can read description of the dataset [HERE](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/unemployment.html).

```{r}
jobless <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/unemployment.csv")
```

Use `glimpse()` to explore the structure of the `jobless` dataset.

```{r}

```

#### Time Series Plot of Unemployment

* Create a line plot for unemployment `rate` (y) by `date` (x).
* Add a filled rectangle to highlight the 1973 to 1975 recession.
* Add a filled rectangle to highlight the 1980 to 1983 recession.
* Add a filled rectangle to highlight the 2007 to 2010 recession.
* Add a filled rectangle to highlight the 2020 to 2021 recession.
* The rectangles should range from 0 to 15 in the y direction.
* Make sure your rectangles do not obscure any other information.
* Add an informative title and appropriate labels for both axes.
* You may optionally add a subtitle or caption for more clarity.
* Be sure to indicate somewhere what the rectangles represent!
* Also be sure to make clear that unemployment rate is a %.
* Add any other formatting you feel will enhance meaning.

```{r}

```

Do you notice any trend(s) in the unemployment rate related to the recession periods? Discuss.

**ANSWER:** type your answer here


### Inflation Dataset

#### Explore the Data

You can read description of the dataset [HERE](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/annualinflation.html).

```{r}
inflate <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/annualinflation.csv")
```

Use `glimpse()` to explore the structure of the `inflate` dataset.

```{r}

```

#### Time Series Plot of Inflation

* Create a time series bar plot for inflation where positive and negative bars are colored differently, as illustrated in Section 3.5 of the [R Graphics Cookbook, 2nd Ed.](https://r-graphics.org/). 
* Use line segment and/or arrow annotation with appropriate text to indicate the span of the Great Depression (1929 to 1939) and the two years with the highest inflation. 
* As always, add an informative title and appropriate labels for both axes, as well as any other formatting you feel will enhance the clarity or meaning if your plot. 
* Be sure a reader can tell from your plot which years are included and which type of inflation value is being displayed (there is more than one way to calculate inflation).

```{r}

```

Discuss two interesting or surprising features of this plot.

**ANSWER:** type your answer here


<hr>

```{r}
sessionInfo()
```

