---
title: "Intro to Reshaping with Separate/Unite and Pivot"
subtitle: "STT2860 Intro to Data Management and Vis S22"
author: "Author: YOUR NAME HERE"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output: 
  html_document: 
    theme: cosmo
    highlight: textmate
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```
<hr>

This worksheet is meant to be used with:

* [RDS CH12 Tidy data](https://r4ds.had.co.nz/tidy-data.html)
* [dplyr: separate and unite](https://youtu.be/ztcsf7G4S78) (video)
* [How to reshape data with tidyr’s new pivot functions](https://youtu.be/gnUgSkKEW5c) (video)
* [Function reference for tidyr](https://tidyr.tidyverse.org/reference/index.html)

## Packages Used

```{r}
library(tidyverse)
library(skimr)
```

## Datasets Used

Example datasets used in **RDS CH12** from the **tidyr** package.

```{r}
table1
table2
table3
table4a
table4b
```

Example dataset from Dr. Thomley's Class Data Files (linked from AsULearn).

```{r}
powerball <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/powerballnumbers.csv")
head(powerball)
```

Example dataset originally from the Kaggle repo [Netflix Movies and TV Shows](https://www.kaggle.com/shivamb/netflix-shows).

```{r}
netflix <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Kaggle/netflix_2021_10_01.csv")
head(netflix)
```

## Separating and Uniting Columns

### `separate()`

```{r}

```

### `unite()`

```{r}

```

## Pivoting the Data "Rectangle"

### `pivot_longer()`

```{r}

```

### `pivot_wider()`

```{r}

```

## Netflix Example

### `separate()` movie categories

Let's first try four categories...

```{r}
netflix_4cat <- netflix %>%
  separate(
    col = listed_in,
    into = c("cat1", "cat2", "cat3", "cat4"),
    sep = ",",
    remove = FALSE
  )

netflix_4cat %>% 
  select(cat1:cat4) %>% 
  skim()
```

We can reduce the number of categories...

```{r}

```

Use `pivot_longer()` to put all the categories in the same column.

```{r}

```

Remove any rows that have `NA` values in the new column you created.

```{r}

```

<hr>
```{r}
sessionInfo()
```
