---
title: "Intro to Data Wrangling with dplyr Verbs"
subtitle: "STT2860 Intro to Data Management & Visualization S23"
author: "Author: YOUR NAME HERE"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output: 
  html_document: 
    theme: cosmo
    highlight: textmate
    toc: true
    toc_depth: 4
    toc_float: true
    code_folding: hide
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```
<hr>

This worksheet is meant to be used with:

* [Intro to Data Wrangling with dplyr Verbs, Part 1](https://stat-jet-asu.github.io/Slides/GeneralR/DataWranglingVerbs01.html#1) slides
* [Intro to Data Wrangling with dplyr Verbs, Part 2](https://stat-jet-asu.github.io/Slides/GeneralR/DataWranglingVerbs02.html#1) slides

## Packages Used

```{r}
library(tidyverse)
library(skimr)
```

## Dataset Used

[Galton Data Description Page](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/galtonheightdata.html)

```{r}
galtondata <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/galtonfamiliesmain.csv")
glimpse(galtondata)
```

## Explore Structure

### `attributes()`

```{r}

```

### `names()` and `row.names()`

```{r}

```

### `dim()`, `nrow()`, and `ncol()`

```{r}

```

### `head()` and `tail()`

```{r}

```

### `skim()` and `skim_without_charts()`

    skim(data, ...)
    skim_without_charts(data, ...)

```{r}

```

```{r}

```

## Data Manipulation

### Using `mutate()`

#### `Mother`

```{r}

```

#### `Child`

```{r}

```

#### `Midparent`

```{r}

```

#### All in One

```{r}

```

### Using `transmute()`

```{r}

```

### Using `pull()`

```{r}

```

### Using `rename()`

```{r}

```

### Using `relocate()`

```{r}

```

### Using `select()`

```{r}

```

## Challenges!

### Challenge #1 

Remove only the variable `Children` from the dataset.

```{r}

```

### Challenge #2

Extract the `HeightT` column from the dataset into a vector.

```{r}

```

### Challenge #3

Convert all height measurements from inches to centimeters.

```{r}

```

### Challenge #4

Create `MotherT` and move it so that it is located after `Mother.`

```{r}

```

### Challenge #5

Start with the original data and rename all the variables with names of your choice.

```{r}

```

### Challenge #6

Start with the original data and create a dataset that contains the following variables: `FamilyID`, `Father`, `MotherT`, `HeightT`, `Child` (in that order).

```{r}

```

### Challenge #7

Explore another use for one of these functions that we have not yet discussed.

```{r}

```


<hr>
```{r}
sessionInfo()
```
